Eurovision Song Contest: semifinal 1, 2009

### 2020-03-05 inst/eurovision.Rmd
euro -> eurovision

The dataset `wiki_matrix`, defined below, is copied from "Eurovision
Song Contest 2009," Wikipedia, accessed May 13, 2018.  It refers to
semi-final 1.  More documentation is given in `euro.Rd` [type
`help(euro2009)` at the `R` prompt].

@@

```{r, specifymatrix}
wiki_matrix <- matrix(c(
#   ME  CZ  BE  BY  SW  AM  AD  CH  TR  IL  BG  IS  MK  RO  FI  PT  MT  BA  DE  UK
    NA, 00, 00, 03, 00, 05, 01, 02, 05, 01, 00, 00, 08, 00, 00, 01, 06, 10, 02, 00, # ME
    00, NA, 00, 00, 00, 00, 00, 00, 00, 00, 00, 00, 00, 00, 00, 00, 00, 00, 00, 00, # CZ
    00, 00, NA, 00, 00, 01, 00, 00, 00, 00, 00, 00, 00, 00, 00, 00, 00, 00, 00, 00, # BE
    02, 01, 00, NA, 01, 04, 00, 00, 00, 04, 01, 01, 06, 00, 04, 00, 01, 00, 00, 00, # BY
    00, 06, 04, 07, NA, 08, 07, 04, 04, 07, 00, 10, 03, 04, 10, 08, 08, 04, 04, 07, # SW
    04, 12, 10, 10, 05, NA, 00, 01, 10, 10, 08, 02, 02, 08, 01, 00, 00, 01, 10, 05, # AM
    00, 00, 00, 00, 00, 00, NA, 00, 01, 00, 00, 00, 00, 00, 00, 04, 03, 00, 00, 00, # AD
    00, 00, 00, 02, 02, 00, 02, NA, 00, 00, 00, 00, 00, 00, 05, 02, 00, 02, 00, 00, # CH
    08, 05, 12, 06, 07, 10, 05, 12, NA, 06, 12, 07, 12, 12, 07, 05, 10, 12, 12, 12, # TR
    05, 04, 03, 04, 06, 07, 08, 05, 03, NA, 04, 06, 01, 03, 06, 00, 04, 00, 05, 01, # IL
    00, 00, 00, 00, 00, 00, 00, 00, 02, 00, NA, 00, 05, 00, 00, 00, 00, 00, 00, 00, # BG
    07, 10, 07, 12, 12, 12, 10, 07, 08, 12, 06, NA, 04, 10, 12, 12, 12, 07, 06, 08, # IS
    10, 03, 00, 00, 00, 00, 00, 06, 06, 00, 10, 00, NA, 02, 00, 00, 00, 08, 00, 00, # MK
    06, 00, 02, 01, 00, 02, 04, 00, 07, 08, 05, 04, 07, NA, 00, 10, 02, 06, 01, 02, # RO
    03, 00, 01, 00, 10, 00, 03, 00, 00, 00, 00, 12, 00, 01, NA, 03, 05, 00, 00, 04, # FI
    00, 02, 06, 00, 03, 00, 12, 10, 00, 02, 02, 08, 00, 07, 02, NA, 00, 03, 07, 06, # PT
    01, 07, 08, 08, 04, 03, 06, 03, 00, 05, 03, 05, 00, 06, 03, 06, NA, 05, 03, 10, # MT
    12, 08, 05, 05, 08, 06, 00, 08, 12, 03, 07, 03, 10, 05, 08, 07, 07, NA, 08, 03) # BA
, nrow=18,byrow=TRUE)
```

Each row corresponds to a competitor and each column to a judge.  Note
that the matrix is not square: the last two columns correspond to
Germany and the UK, who were judges but not competitors.  I have
removed the first column in the Wikipedia table (which gave the total
of the points), replaced self-voting entries with `NA`, and replaced
blanks with `00`; every entry is padded to two digits.  Thus the first
row corresponds to votes cast for `ME`, Montenegro.  We see that `BE`
(Belgium) awarded them three points, `am` (Armenia) five points, and
so on.

@@

Now, take the first row of `preference`.  This represents the votes
cast *by* (sic) Montenegro (`"ME"`).  Their favourite was [last
column] Bosnia & Herzegovina (`"BA"`), who they gave rank 1 to.  Their
second favourite was Macedonia, their third was Turkey, and so on.
They were not allowed to vote for themselves, which is why the first
column is `NA`.  So the order was: `BA` (1), `MK` (2) , `TR` (3), `IS`
(4), `RO` (5), `IL` (6), `AM` (7), `FI` (8), `BY` (9), `MT` (10).  The
other countries (`CZ`, `BE`, `SW`, `CH`, `BG`, `PT`) did not receive
any points from Montenegro.  We need to convert matrix `preference`
into a likelihood function, here `euro2009`:

```{r,morestuff}
euro2009 <- hyper2(d=18)

for(i in seq_len(nrow(preference))){   # cycle through the rows; each row is a voter
    d <- preference[i,,drop=TRUE]
    d[is.na(d)] <- -1  # kludge: make the voting country ineligible to vote.
    while(any(d>0)){
        eligible <- which(d>=0)   # This is why we set NA values to -1
        euro2009[which(d==1)] %<>% "+"(1)
                                        # The first choice among
                                        # eligible players has +1
                                        # power on the numerator

        euro2009[eligible] %<>% dec()   # denominator of all eligible players

@@

## syntatic sugar:
pnames(euro2009) <- competitors
```

@@

```{r,checknull}
m <- maxp(euro2009)  # MLE
print(loglik(indep(m),euro2009))
print(loglik(indep(equalp(euro2009)),euro2009))
```

>>>>

The dataset `wiki_matrix`, defined below, is copied from "Eurovision
Song Contest 2009," Wikipedia, accessed May 13, 2018.  It refers to
semi-final 1.  More documentation is given in `eurovision.Rd` [type
`help(euro2009)` at the `R` prompt].

@@

```{r, specifymatrix}
data(eurovision)
(wiki_matrix <- as.matrix(eurovision2009_votingtable))
```

Each row corresponds to a competitor and each column to a judge.  Note
that the matrix is not square: the last two columns correspond to
Germany and the UK, who were judges but not competitors.  I have
removed the first column in the Wikipedia table (which gave the total
of the points), replaced self-voting entries with `NA`, and replaced
blanks with `0`.  Thus the first row corresponds to votes cast for
`ME`, Montenegro.  We see that `BE` (Belgium) awarded them three
points, `am` (Armenia) five points, and so on.

@@

Now, take the first row of `preference`.  This represents the votes
cast *by* (sic) Montenegro (`"ME"`).  Their favourite was [last
column] Bosnia & Herzegovina (`"BA"`), who they gave rank 1 to.  Their
second favourite was Macedonia, their third was Turkey, and so on.
They were not allowed to vote for themselves, which is why the first
column is `NA`.  So the order was: `BA` (1), `MK` (2) , `TR` (3), `IS`
(4), `RO` (5), `IL` (6), `AM` (7), `FI` (8), `BY` (9), `MT` (10).  The
other countries (`CZ`, `BE`, `SW`, `CH`, `BG`, `PT`) did not receive
any points from Montenegro.  We need to convert matrix `preference`
into a likelihood function, here `eurovision2009`:

```{r,morestuff}
eurovision2009 <- hyper2(d=18)

for(i in seq_len(nrow(preference))){   # cycle through the rows; each row is a voter
    d <- preference[i,,drop=TRUE]
    d[is.na(d)] <- -1  # kludge: make the voting country ineligible to vote.
    while(any(d>0)){
        eligible <- which(d>=0)   # This is why we set NA values to -1
        eurovision2009[which(d==1)] %<>% "+"(1)
                                            # The first choice among
                                             # eligible players has +1
                                              # power on the numerator

        eurovision2009[eligible] %<>% dec()   # denominator of all eligible players

@@

## syntatic sugar:
pnames(eurovision2009) <- competitors
```

@@

```{r,checknull}
m <- maxp(eurovision2009)  # MLE
print(loglik(indep(m),eurovision2009))
print(loglik(indep(equalp(eurovision2009)),eurovision2009))
```
